\documentclass{article}
\usepackage{bubblecv}
\usepackage{graphicx} % Required for inserting images
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{minted}
\usepackage{textcomp}
\usepackage{mathtools}
\usepackage{float}
% Set the paragraph spacing
\setlength{\parskip}{0.5\baselineskip} % Set space between paragraphs
\setlength{\parindent}{0pt} % Optionally set the paragraph indentation to 0
\begin{document}



Application for the Position Risk Analyst at
Norsk Hydro ASA.

Hi, my name is Filmon. Within the next two months I will be completing my master's degree in Mathematics for Applications at the University of Oslo. My thesis focuses on developing numerical methods for solving partial differential equations (PDEs), applied to traffic simulation.

The work involves extensive numerical analysis and a significant amount of programming in Python. I have to be systematic when understanding the complex model and designing numerical algorithms. I also make sure to write structured code and document results clearly so they are easy for others to understand. This experience has given me solid skills in data analysis, model development, and programming. I am therefore motivated to join year team as Risk Analyst. I believe this role will offer me the opportunity to learn and grow.

I am applying for this position because I want to keep my skills in mathematics and programming up to date with new technology. In addition, I have a strong desire to expand my network. I believe that this role will give me valuable experience and a strong foundation for further development.

I have strong programming skills. I can program in Python, and I am comfortable using GitHub. I enjoy collaborating on project based work. You can find the implementation of my master project at the following GitHub link:
https://github.com/Filmon492/MAT5960

With my background in mathematics and programming, I believe I will adapt quickly and contribute effectively to your team. I enjoy challenging others by asking good questions, and I am always interested in exploring new methods. I want to fully understand why one method works better than another. I look forward to the opportunity to work with you.

Kind regards,\\
Filmon

%I believe that my ability to build mathematical models and validate them to real-word scenario would be significant for this position. Special courses such as spline methods, the Finite Element Method (FEM) and the Finite Volume Method (CFD) have provided me solid computational skills, which enabling me to tackle complex problems at your company.

%I have experience working in a team. I work together with another person as a mentor at ENT3R UiO. We help high school students with math. We communicate well and share the tasks equally between us. This experience will help me to collaborate at your company.


%With my mathematical and programming skills, I believe I will adapt quickly and achieve competence in your team. I enjoy challenging others by asking good questions. I am always interested in exploring new methods and want to understand completely why one method works better than others. I look forward to hearing from you.





\end{document}